### 2023-05-16 definitions.tex
sections

The model we are interested in is labeled WSTS (and later some particular instances, i.e. Timed/Counter Automata for instance).

@@

\noindent
{\bf Transition systems}

\begin{definition}
 A {\em transition system} is a pair $\mathscr{S} = (S,\rightarrow )$ where $S$ is a set of 
 {\em states} and  
 $ {\rightarrow} \subseteq S \times S$ is a
 binary relation 
 on
 the set of states, denoted as the set of {\em transitions}.

\iffalse
\mathieu{ Maybe add this back ?
A {\em labeled transition system} (LTS for short) is a tuple $\mathscr{S}=(S, \Sigma, \rightarrow)$ where 
$S$ is a set of {\em states}, $\Sigma$ is a (finite) set of {\em labels}, and 
${\rightarrow} \subseteq S\times \Sigma \times S$ is a 
ternary relation,
denoted as the set of {\em labeled transitions}. }
\fi
\end{definition}

% We prefer to use infix notation and $(s,a ,s')\in {\rightarrow} $ will be abbreviated as    $s  \xrightarrow{a}  s'$ to represent a transition from configuration $s$ to configuration $s'$ with label $a$. 
We prefer to use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from state $s$ to state $s'$ (i.e., $ (s,s') \in  {\rightarrow} $).
We write $\rightarrow^{k}$, $\rightarrow^{+}$, $\rightarrow^{=}$, $\rightarrow^{*}$
for the $k$-step iteration of $\rightarrow$, its transitive closure, its reflexive closure, its reflexive and transitive closure.
Let $X,Y \subseteq S$ and $k \in \mathbb{N}$. We denote $X \longrightarrow^{*} Y$ (resp. $X \longrightarrow^{\leq k} Y$) if from all states $x \in X$ there exists a path (resp. of length smaller than $k$) that reaches a state $y \in Y$.
\noindent
The set of {\em (immediate) successors} of a state $s \in S$ is defined as 
 $\post(s) = \{ s' \in S \mid  ~ s \xrightarrow{} s'\}$. 
% $\post(s) = \{ s' \in S \mid  ~ \exists a \in \Sigma~ s \xrightarrow{a} s'\}$. 
A state without successors is called a {\em dead end}. 
The set of {\em (immediate) predecessors} of a state $s \in S$ is deﬁned as
 $\pred(s) = \{ s' \in S \mid  ~ s' \xrightarrow{} s\}$. 
% $\pred(s) = \{ s' \in S \mid  ~ \exists a \in \Sigma~ s' \xrightarrow{a} s\}$. 
By iterating $\pred$ and $\post$ we obtain  
$\post^n(s) = \{ s' \in S \mid  ~ s' \xrightarrow{}^n s\}$
and
$\pred^n(s) = \{ s' \in S \mid  ~ s' \xrightarrow{}^n s\}$.
However, we are generally more interested in
$\post^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \post^i(s)$, $\post^*(s)= \bigcup_{1 \leq i} \post^i(s)$
and
$\pred^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \pred^i(s)$ and $\pred^*(s) = \bigcup_{1 \leq i} \pred^i(s)$. % \\

@@

\noindent
{\bf Well-quasi-orderings}

A {\em quasi-ordering} (a qo) is any reflexive and transitive relation $\leq$ over some set $X$ and we often write $(X,\leq)$.

We abbreviate $x \leq y \not\leq x$ by $x < y$.
% A partial ordering (a po) is an antisymmetric qo
	% \alain{
Any qo induces an equivalence relation ($x \equiv y$ iff $x \leq y \leq x$).
	% }
% and gives rise to a po between the equivalence classes ?
% 	do we need po ?

We now recall a few results from the theory of well-orderings (add reference [...]).

\begin{definition}
 A {\em well-quasi-ordering} (wqo) is any quasi-ordering $(X,\leq)$ such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with $x_i \leq  x_j$.
\end{definition}

There exists a stronger formulation. $(X,\leq)$ is a wqo iff any infinite sequence  $x_0, x_1, x_2, ...$ in $X$ contains an infinite increasing subsequence: $x_{i_0} \leq x_{i_1} \leq x_{i_2} \cdots$ (with $i_0 < i_1 < i_2 \cdots$).

% \textcolor{red}{Add lemma about infinite increasing subsequences ?}
%
\iffalse
%
\begin{lemma}
(Erd\"os and Rado). Assume $\leq$ is a wqo. Then any infinite sequence contains an infinite increasing subsequence: $x_{i_0} \leq x_{i_1} \leq x_{i_2} \cdots$ (with $i_0 < i_1 < i_2 \cdots$).
\end{lemma}

@@

Given $(X,\leq)$ a quasi-ordering, an {\em upward-closed set} is any set $U \subseteq X$ such that if $y \geq x$ and $x \in U$ then $y \in U $.
A {\em downward-closed set} is any set $D \subseteq X$ such that if $y \leq x$ and $x \in D$ then $y \in D $. 
It is an {\em ideal } if it is also {\em directed}, i.e. it is nonempty and for every $a,b \in D$, there exists $c \in D$ such that $a \leq c$ and $b \leq c$.
To any subset $A \subseteq X$, we may associate
its {\em upward-closure},
 $\uparrow A = \{x \in X \mid \exists a \in A ~ y \geq a\}$
 and its 
 {\em downward-closure},
 $\downarrow A = \{x \in X \mid \exists a \in A ~ y \leq a\}$. 
We abbreviate $\uparrow \{x\}$ (resp. $\downarrow \{x\}$)
as $\uparrow x$ (resp. $\downarrow x$).

A {\em basis} of an upward-closed set $I$ is a set $I_b$ such that $I = \uparrow I_b$. 
% Higman investigated ordered sets with the finite basis property.

Wqo admits many other equivalent formulations. For instance: 
$(X,\leq)$ is a wqo if and only if any infinite increasing sequence $U_0 \subseteq U_1 \subseteq U_2 \subseteq \cdots$ of
upward-closed subsets in $X$ eventually stabilizes, 
%      i.e. there is a $k \in N$ such that $U_k = U_{k+1} = U_{k+2} = \cdots $ 
if and only if any upward-closed subset $U \subseteq X$ has a
finite basis if and only if $(X,\leq)$ is well-founded (i.e. it admits no infinite strictly decreasing sequence) and $X$ don't contain infinite subset of mutually incomparable elements (antichains).
%
\iffalse
%

@@

Quasi-orderings that have no infinite antichains enjoy a similar \emph{finite decomposition} property than wqo. In such \emph{Finite Antichains} sets $(X,\leq)$, every downward closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2,..., J_n$ (an ideal is a directed downward closed subset) such that $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.

@@

\noindent
{\bf Well-structured transition systems}

@@

A WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ is {\em effective} if there exists a pair of Turing machines
($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
$ M_\rightarrow$ computes the transition relation “$\rightarrow$” and 
$M_\leq$ the ordering relation “$\leq$”. 
We say that $\mathscr{S}$ is {\em post-effective} if it is effective, and if there
exists an additional Turing machine that computes $|\post(x)| \in \N \cup \{\infty\}$ 
on input $x$,
with $x \in X_i $. 
Such a Turing machine, in combination with $M_\rightarrow$, allows computing
$\post(x)$ whenever the latter is finite. We say that $\mathscr{S}$ is {\em upward pre-effective} if it is effective,
and if there exists an additional Turing machine that computes a finite basis of 
$\uparrow \pred (\uparrow x)$
on input $x$, where $x$ is a state of $\mathscr{S}$. 
We say that $\mathscr{S}$
is {\em ideally effective} if
\begin{itemize}
\item the function mapping the encoding of a configuration $s$
to the encoding of the ideal $\downarrow s$ is computable;
\item inclusion of ideals is decidable;
\item the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
\end{itemize}

\iffalse
\alain{pas ici}
\begin{definition}{\cite{BFM-ic17}} 
A class $C$ of WSTS is {\em ideally effective} if
\begin{itemize}
\item the function mapping the encoding of a configuration $s$ of a WSTS
to the encoding of the ideal $\downarrow s$ is computable;
\item inclusion of ideals is decidable;
\item the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
\end{itemize}
\end{definition}

\fi

We will use the following property: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
%\alain{étrange ces deux propositions, l'une suffirait}
\iffalse
Proof. Assume $s \in \pred (I )$. Then $s \rightarrow t$ for some $t \in I $. If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$. Then $t' \in I$ and $s' \in \pred(I )$.
\fi
%
Now, we recall a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.

@@

@@

\begin{theorem}(ref)
A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ with effective pred-basis and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$.
\end{theorem}

@@

% (ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
% \alain{(ii) est bizarre car A est déjà une base finie de $\uparrow A$, et $\min(A)$ est la base minimale}
%			\end{fact}
For a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and an upward-closed set $U  \subseteq S$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U_{k-1} \cup \pred(U_{k-1})$ for $k \geq 1$. When $\mathscr{S}$ has the strong compatibility, the sets $U_k$ are upward-closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)_k$ converges. Let us define the \emph{index} of convergence of the sequence $U_k$ as the smallest $k_0$ s.t. $U_k = U_{k_0}$ for all $k \geq k_0$. We may compute $k_0$ and we then have:  $\pred^*(U) = U_{k_0}$. \\
%

>>>>

%The model we are interested in is labeled WSTS (and later some particular instances, i.e. Timed/Counter Automata for instance).

@@

\noindent
Let us recall thz well-known framework of transition systems. 
 A {\em transition system} is a pair $\mathscr{S} = (S,\rightarrow )$ where $S$ is a set of 
 {\em states} and  
 $ {\rightarrow} \subseteq S \times S$ is a
 binary relation 
 on
 the set of states, denoted as the set of {\em transitions}. 
%
\iffalse
\mathieu{ Maybe add this back ?
A {\em labeled transition system} (LTS for short) is a tuple $\mathscr{S}=(S, \Sigma, \rightarrow)$ where 
$S$ is a set of {\em states}, $\Sigma$ is a (finite) set of {\em labels}, and 
${\rightarrow} \subseteq S\times \Sigma \times S$ is a 
ternary relation,
denoted as the set of {\em labeled transitions}. }
\fi
%
% We prefer to use infix notation and $(s,a ,s')\in {\rightarrow} $ will be abbreviated as    $s  \xrightarrow{a}  s'$ to represent a transition from configuration $s$ to configuration $s'$ with label $a$. 
We use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from state $s$ to state $s'$ (i.e., $ (s,s') \in  {\rightarrow} $).
We write $\rightarrow^{k}$, $\rightarrow^{+}$, $\rightarrow^{=}$, $\rightarrow^{*}$
for the $k$-step iteration of $\rightarrow$, its transitive closure, its reflexive closure, its reflexive and transitive closure.
Let $X,Y \subseteq S$ and $k \in \mathbb{N}$. We denote $X \longrightarrow^{*} Y$ (resp. $X \longrightarrow^{\leq k} Y$) if from all states $x \in X$ there exists a path (resp. of length smaller than $k$) that reaches a state $y \in Y$.
\noindent
The set of {\em (immediate) successors} of a state $s \in S$ is defined as 
 $\post(s) = \{ s' \in S \mid  ~ s \xrightarrow{} s'\}$. 
% $\post(s) = \{ s' \in S \mid  ~ \exists a \in \Sigma~ s \xrightarrow{a} s'\}$. 
A state without successors is called a {\em dead end}. 
The set of {\em (immediate) predecessors} of a state $s \in S$ is deﬁned as
 $\pred(s) = \{ s' \in S \mid  ~ s' \xrightarrow{} s\}$. 
% $\pred(s) = \{ s' \in S \mid  ~ \exists a \in \Sigma~ s' \xrightarrow{a} s\}$. 
By iterating $\pred$ and $\post$ we obtain  
$\post^n(s) = \{ s' \in S \mid  ~ s' \xrightarrow{}^n s\}$
and
$\pred^n(s) = \{ s' \in S \mid  ~ s' \xrightarrow{}^n s\}$.
However, we are generally more interested in
$\post^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \post^i(s)$, $\post^*(s)= \bigcup_{1 \leq i} \post^i(s)$
and
$\pred^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \pred^i(s)$ and $\pred^*(s) = \bigcup_{1 \leq i} \pred^i(s)$. % \\

@@

\noindent
A {\em quasi-ordering} (a qo) is any reflexive and transitive relation $\leq$ over some set $X$ and we often write $(X,\leq)$. We abbreviate $x \leq y \not\leq x$ by $x < y$.
% A partial ordering (a po) is an antisymmetric qo
	% \alain{
Any qo induces an equivalence relation ($x \equiv y$ iff $x \leq y \leq x$).
	% }
% and gives rise to a po between the equivalence classes ?
% 	do we need po ?
Given $(X,\leq)$ a quasi-ordering, an {\em upward-closed set} is any set $U \subseteq X$ such that if $y \geq x$ and $x \in U$ then $y \in U $.
A {\em downward-closed set} is any set $D \subseteq X$ such that if $y \leq x$ and $x \in D$ then $y \in D $. 
It is an {\em ideal } if it is also {\em directed}, i.e. it is nonempty and for every $a,b \in D$, there exists $c \in D$ such that $a \leq c$ and $b \leq c$.
To any subset $A \subseteq X$, we may associate
its {\em upward-closure},
 $\uparrow A = \{x \in X \mid \exists a \in A ~ y \geq a\}$
 and its 
 {\em downward-closure},
 $\downarrow A = \{x \in X \mid \exists a \in A ~ y \leq a\}$. 
We abbreviate $\uparrow \{x\}$ (resp. $\downarrow \{x\}$)
as $\uparrow x$ (resp. $\downarrow x$).
%
A {\em basis} of an upward-closed set $I$ is a set $I_b$ such that $I = \uparrow I_b$.

\begin{definition}
 A {\em well-quasi-ordering} (wqo) is any quasi-ordering $(X,\leq)$ such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with $x_i \leq  x_j$.
\end{definition}
%
% \textcolor{red}{Add lemma about infinite increasing subsequences ?}
%
\iffalse
%
\begin{lemma}
(Erd\"os and Rado). Assume $\leq$ is a wqo. Then any infinite sequence contains an infinite increasing subsequence: $x_{i_0} \leq x_{i_1} \leq x_{i_2} \cdots$ (with $i_0 < i_1 < i_2 \cdots$).
\end{lemma}

@@

% Higman investigated ordered sets with the finite basis property.

Wqo admits many other equivalent formulations. For instance: 
$(X,\leq)$ is a wqo f and only if any infinite sequence  $x_0, x_1, x_2, ...$ in $X$ contains an infinite increasing subsequence: $x_{i_0} \leq x_{i_1} \leq x_{i_2} \cdots$ (with $i_0 < i_1 < i_2 \cdots$)
if and only if any infinite increasing sequence $U_0 \subseteq U_1 \subseteq U_2 \subseteq \cdots$ of
upward-closed subsets in $X$ eventually stabilizes, 
%      i.e. there is a $k \in N$ such that $U_k = U_{k+1} = U_{k+2} = \cdots $ 
if and only if any upward-closed subset $U \subseteq X$ has a
finite basis if and only if $(X,\leq)$ is well-founded (i.e. it admits no infinite strictly decreasing sequence) and $X$ don't contain infinite subset of mutually incomparable elements (antichains).
%
\iffalse
%

@@

\noindent
Quasi-orderings that have no infinite antichains enjoy a similar \emph{finite decomposition} property than wqo. In such \emph{Finite Antichains} sets $(X,\leq)$, every downward closed subset $D \subseteq X$ can be decomposed into a \emph{finite} set of ideals $J_1, J_2,..., J_n$ (an ideal is a directed downward closed subset) such that $D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$.

@@

\noindent
Let us now recall the most general definition of well-structured transition systems.

@@

A WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ is {\em effective} if there exists a pair of Turing machines
($M_\rightarrow$, $M_\leq$) operating on $\N \times \N$ such that
$ M_\rightarrow$ computes the transition relation “$\rightarrow$” and 
$M_\leq$ the ordering relation “$\leq$”.

We say that $\mathscr{S}$ is {\em post-effective} if it is effective, and if there
exists an additional Turing machine that computes $|\post(x)| \in \N \cup \{\infty\}$ 
on input $x$, with $x \in X_i $. 
Such a Turing machine, in combination with $M_\rightarrow$, allows computing
$\post(x)$ whenever the latter is finite.

We will use the following property: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are all upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
%\alain{étrange ces deux propositions, l'une suffirait}

\iffalse
Proof. Assume $s \in \pred (I )$. Then $s \rightarrow t$ for some $t \in I $. If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$. Then $t' \in I$ and $s' \in \pred(I )$.
\fi
%

Now, we recall a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.

@@

For a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and an upward-closed set $U  \subseteq S$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U_{k-1} \cup \pred(U_{k-1})$ for $k \geq 1$. When $\mathscr{S}$ has the strong compatibility, the sets $U_k$ are upward-closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)_k$ converges. Let us define the \emph{index} of convergence of the sequence $U_k$ as the smallest $k_0$ s.t. $U_k = U_{k_0}$ for all $k \geq k_0$. We may compute $k_0$ and we then have:  $\pred^*(U) = U_{k_0}$. \\

@@

\begin{theorem}\cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00}
A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ with effective pred-basis and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$. Hence coverability is decidable.
\end{theorem}

%We say that $\mathscr{S}$ is {\em upward pre-effective} if it is effective,
%and if there exists an additional Turing machine that computes a finite basis of 
%$\uparrow \pred (\uparrow x)$
%on input $x$, where $x$ is a state of $\mathscr{S}$.

We say that $\mathscr{S}$
is {\em ideally effective} \cite{BFM-ic17} if (1) the function mapping the encoding of a configuration $s$
to the encoding of the ideal $\downarrow s$ is computable; (2) inclusion of ideals is decidable; (3) the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
%
\iffalse
\alain{pas ici}
\begin{definition}{\cite{BFM-ic17}} 
A class $C$ of WSTS is {\em ideally effective} if
\begin{itemize}
\item the function mapping the encoding of a configuration $s$ of a WSTS
to the encoding of the ideal $\downarrow s$ is computable;
\item inclusion of ideals is decidable;
\item the downward closure $\downarrow \post(I)$ expressed as a finite union of ideals is computable from the ideal $I$.
\end{itemize}
\end{definition}
%
\fi
%
%
With the ideally effective hypothese, we obtain the decidability of coverability for a class of ordered transition systems larger than WSTS.

\begin{theorem}\cite{BFM-ic17}
Coverability is decidable for any ideally effective ordered compatible transition system $\mathscr{S}=(S, \rightarrow, \leq)$ where $\leq$ is without infinite antichains.
\end{theorem}

@@

% (ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
% \alain{(ii) est bizarre car A est déjà une base finie de $\uparrow A$, et $\min(A)$ est la base minimale}
%			\end{fact}
%

### 2023-05-16 introduction.tex
up
sections

{\bf Context.} In the quite informal paper \cite{DBLP:journals/corr/PrasadZ16}, Prasad and Zuck studied resilience of a process to an adversary and they used the framework of WSTS enjoying both upward and downward monotonies to prove the decidability of a kind of resilience.

{\bf MFCS 2023
Abstract submission deadline:    	 April 24th (AoE) 
Paper submission deadline:    	April 28th (AoE)} \\
%
{\bf CONCUR 2023 Abstract Registration Due 	Apr 24, 2023
Submission Deadline 	May 2, 2023}\\
%
{\bf FSTTCS 2033  Submission deadline: July 14, 2022 AoE (firm)
Notification to authors: September 16, 2022}

@@

\begin{center}
	\begin{figure}
			\hspace{0.75cm}
				\label{air control}
\includegraphics[width=0.85\textwidth]{FigureB}
	\caption{A channel system with three automata and four channels. Leftmost and rightmost automata represents aircrafts attempting to land and staying on stand-by in the air while they haven't received permission to land, while the middle one represent the control tower managing the aircrafts landings so that only one aircraft at a time is allowed for landing.}
	\end{figure}
\end{center}

@@

>>>>

{\bf Context.} In the quite informal paper \cite{DBLP:journals/corr/PrasadZ16}, Prasad and Zuck studied resilience of a process to an adversary and they used the framework of WSTS enjoying both upward and downward monotonies to prove the decidability of a kind of resilience.
%
%{\bf MFCS 2023
%Abstract submission deadline:    	 April 24th (AoE) 
%Paper submission deadline:    	April 28th (AoE)} \\
%%
%{\bf CONCUR 2023 Abstract Registration Due 	Apr 24, 2023
%Submission Deadline 	May 2, 2023}\\
%
{\bf FSTTCS 2033  Submission deadline: July 14, 2022 AoE (firm)
Notification to authors: September 16, 2022}

@@

Resilience is an instance of the home space problem: let $L_1,L_2$ be two sets of states, the home space for  $L_1,L_2$ is the following question:		
	$\post^*(L_1)	\subseteq \pred^*(L_2)$
\alain{ne trouve pas la macro pred et post}

\begin{center}
	\begin{figure}
			\hspace{0.75cm}
\includegraphics[width=0.85\textwidth]{FigureB}
	\caption{A channel system with three automata and four channels. Leftmost and rightmost automata represents aircrafts attempting to land and staying on stand-by in the air while they haven't received permission to land, while the middle one represent the control tower managing the aircrafts landings so that only one aircraft at a time is allowed for landing.}
					\label{air control}
	\end{figure}
\end{center}

@@

\alain{il faut mieux expliquer le modèle, le rôle des 4 canaux, des messages, remarquer que 2 canaux sont des compteurs,....}
